Case-based Arabic Morphological Analysis

### 2010-12-17 tex/sarf.aclhlt.tex
results and conclusion

In section we present results that compare Sarf to existing 
morphological analyzers such as Buckwalter and ElixirFM. 
We also evaluate our hypothesis of the utility of a case-based 
morphological analyzer,
by running our case study in two modes, simple and refined.
We ran our experiments using a dual core 2.66 Ghz 64-bit processor 
with 4GB of memory using a Linux operating system.

In the simple mode, Sarf is controlled by the hadith controller.
In the refined mode, we apply several refinements to Sarf.
The first refinement ignore female proper names except at the end of
a chain since they do not occur often in narrator names in hadith while
they can be common in early narrators such as the wives of the prophet.
We also add a refinement that detects compound proper names 
such as \RL{`bd alr.hmn} 
as they are common in narrator names.
Finally, we consider a possessive form as a reference to a person and treat
it as a narrator.

We report our results in Table~\ref{t:hadithresallresults} on the three hadith books.
The word count row is an indicator of the size of data.
The number of names, narrators, and chains reflect a hierarchical structure where a chain contains
narrators, and a narrator contains names.
In general, we observe that the refinements slightly improved performance and accuracy at
the lowest level of abstraction by reducing false positives.
However, the refinements had little to no effect on the high accuracy measures at higher
levels of abstraction.
We believe that this is a strong supporting evidence to our hypothesis.

The names row reports the total number of detected names. The names/narrator row reports the average
number of names per narrator structure. The narrators row reports the number of narrators detected.
The narrators/chain row reports the number of narrators per chain structure. And the chain row reports
the number of detected chains in the book.
The ignored names row reports the number of names detected outside the chains and thus irrelevant to
the case study.
The segmentation accuracy reports on the accuracy of segmentation of the book into separate hadiths.
The metric is computed manually over a representative sample of each text. The same applies to all other accuracy measures.
The chain accuracy reflects the percentage of correct narrators covered by extracted chains.
The narrator name accuracy row reflects the percentage of correct proper names detected within
narrator structures.

The name false positives row reports the percentage of words recognized as names while they should not.
We notice that the low level metric of name false positives improved dramatically with the refined
analysis without substantially affecting the high abstract measures such as chain and segmentation
accuracy.
We notice that the controller in the simple analysis ignored more names and thus the case-based
controller compensated for the inaccuracy of the low level analysis with a simple heuristic and
without the need of the fancy refinements.

We investigated the case of Al Kafi where the accuracy of the simple analysis was significantly lower than that of the refined
analysis. We  discovered that the chains detected by the refined analysis were a subset of those detected by the simple
analysis. 
The reason for the lower accuracy is false positive chains. 
The simple analysis did not miss important information. 
We hypothesize that a higher level of abstraction such as intersecting the narrators from the hadith books with the narrators 
in the biographies can reduce the false positives in the chains.

@@

We implemented our case study with Buckwalter in two phases. 
In phase one, we process the entire text with Buckwalter and
obtain annotations to locations in text. 
Then in the second phase we feed these annotations to 
our controller as if they were coming from Sarf. 
We measure the time it took to produce the annotations
and the time it took the controller to process the 
annotations whithout the time spent streaming the annotations.
We used the same setup for ElixirFM.

Sarf outperforms both Buckwalter and ElixirFM in runing time. 
It also produces higher accuracy measures. 
We think this is mainly due to the extended
lexicon as well as to the fact that Sarf reduces false positive
names by ignoring prefixes and suffixes that can not form
names even if the stem was a name. 
Doing the same with Buckwalter and ElixirFM requires  
expanding the annotations resulting from one affix to reflect
its multiple composing affixes.
This is easily done with the way we defined affixes recursively.

Buckwalter performed much better than ElixirFM. This is up to 
our knowledge the first comparison of these two tools and 
we think the big difference is due to the more exact 
and grammar based functional morphological analysis ElixirFM 
uses.

@@

In the future, we plan to complete the hadith case study 
and automate narrator discovery and authentication in the 
biography books. 
We plan to make use of the name connector and narrator
connector stop phrases that our case study extracted to create a 
learning system that can learn more names. 
We also plan on developing a domain specific language similar 
to ElexirFM~\cite{Otakar:07}
that allows users to easily specify their case-based controller
to interact with Sarf.

>>>>

In this section we present results that compare Sarf to existing 
morphological analyzers such as Buckwalter and ElixirFM. 
We also evaluate our hypothesis of the utility of a case-based 
morphological analyzer.
We validate that the high-level analysis of the output 
of low-level morphological analysis can compensate for accuracy.
To acheive that we ran Sarf in two modes, simple and refined.
In the simple mode, we used Sarf as discussed in 
Section~\ref{sec:sarf} with the 
controller of Section~\ref{sec:controller}.
In the refined mode, we applied several low-level morphological 
refinements to Sarf such as disambiguation of 
female proper names so that the controller can ignore them. 
We also added a refinement that detected compound proper names 
such as \RL{`bd alr.hmn} 
as they are common in narrator names.
%Finally, we consider a possessive form as a reference to 
%a person and treat
%it as a narrator.

We ran our experiments using a dual core 2.66 Ghz 64-bit processor 
with 4GB of memory using a Linux operating system.

We report our results in Table~\ref{t:hadithresallresults} on the 
three hadith books.
In general, we observe that the refinements slightly 
improved performance and accuracy at
the lowest level of abstraction by reducing false positives.
However, the refinements had little to no effect on the high 
accuracy measures at higher levels of abstraction.
We believe that this is a strong supporting evidence to our 
hypothesis.

The row for ``word count'' is an indicator of the size of the data.
The %numbers in 
rows ``names'', ``narrators'', and ``chains'' denote the total
number of the corresponding entity per book.
The relative magnitudes of these rows
reflect the hierarchical structure where 
a chain contains
narrators, and a narrator contains names.
The row ``ignored names'' reports the number of names detected 
outside the chains and thus are irrelevant to
the case study.
In addition, the rows ``names/narrator'' and ``narrators/chain'' 
report the average
number of names per narrator and 
the average number of narrators per chain respectively.

The segmentation accuracy reports on the accuracy of 
segmentation of the book into separate hadiths.
The metric is computed manually over a representative sample 
of each book.
The same applies to all other accuracy measures.
The chain accuracy reflects the percentage of correct narrators 
covered by extracted chains.
The ``narrator name accuracy'' row reflects the percentage of 
correct proper names detected within
narrator structures.

The row ``name false positives'' reports the percentage of words 
recognized as names while they should not.
We notice that this low level metric improved dramatically 
with the refined analysis without substantially affecting the 
high abstract measures such as chain and segmentation
accuracy.

We notice that the controller in the simple analysis ignored 
more names and thus the case-based
controller compensated with a simple heuristic
for the inaccuracy of the low level 
analysis.
%and without the need of the low-level refinements.

We investigated the case of Al-Kafi where the accuracy of 
the simple analysis was significantly lower than that of the 
refined analysis. 
We discovered that the chains detected by the refined analysis 
were a subset of those detected by the simple
analysis. 
The reason for the lower accuracy is false positive chains, 
implying that the simple analysis did not miss important 
information.

We hypothesize that a higher level of abstraction such as 
intersecting the narrators from the hadith books with the narrators 
in the biographies can reduce the false positives in the chains
in this case.

@@

We implemented our case study with Buckwalter in two phases. 
In phase one, we processed the entire text with Buckwalter and
obtained annotations to text. 
In the second phase, we fed the annotations to 
the hadith controller as if they were coming from Sarf. 
We measured the time it took to produce the annotations
and the time it took the controller to process the 
annotations whithout the time spent streaming the annotations.
We used the same setup for ElixirFM.

Sarf outperforms both Buckwalter and ElixirFM in terms of 
running time. 
It also produces higher accuracy measures. 
We think this is mainly due to the extended
lexicon as well as to the fact that Sarf reduces false positive
names by ignoring prefixes and suffixes that can not form
names even if one possible alternative of the stem is a name. 
Doing the same with Buckwalter and ElixirFM requires  
expanding the annotations resulting from one simple affix to be 
reflected in all the composite affixes that contain it.
%expanding the annotations resulting from one affix to reflect
%its multiple composing affixes.
%However, this was easily performed using our recursive definition 
%of the affixes.

Buckwalter performed much better than ElixirFM in 
terms of running time, 
while ElixirFM outperformed Buckwalter in terms of accuracy.
%This is up to 
%our knowledge the first comparison of these two tools and 
We think the big difference is due to the sofistication of 
grammar based functional morphological analysis ElixirFM 
uses and its reliance on additional annotations extracted from both ATB~\cite{Maamouri:04} and the Pragbank~\cite{Prague04}.

@@

We presented Sarf, a novel case-based morphological analyzer,
and validated its utility using the hadith literature
case study. 
Sarf uses non-deterministic finite state Automata with 
concatenative analysis and recursive affixes to analyze 
Arabic text. 
Our results showed that Sarf with the hadith controller
is faster and more accurate than
current morphological analyzers.

In the future, 
we will try Sarf with other interesting case studies.
We plan to complete the hadith case study 
and automate narrator discovery and authentication in the 
biography books. 
We plan to make use of the name connector and narrator
connector stop phrases that our case study extracted to create a 
learning system that can learn more names. 
We also plan on developing an Arabic linguistic computational model 
similar to that of ElixirFM~\cite{Otakar:07} but that can be 
explored partialy based on the case study.
